---
knit: "bookdown::render_book"
---

# A MC for Brazilian Portuguese

## Description

???


## Read the dictionary

* We want to define the `wordlist` variable, which will be a list of character vectors.

* Each vector will contain the characters of a word, with added '^' at the beginning and added double '$' at the end.

* But because the `markovchain` package gets confused by state names containing accented characters, first we will preprocess the dictionary, converting accented characters to pairs of characters:

  * 'á' becomes 'a´'
  * 'ç' becomes 'c,'
  * and so on

* The following awk code will read the dictionary and produce an R script to define the variable `wordlist`:

    ```{awk, engine.opts = './data/brazilian', code = readLines('./data/process.awk'), cache = TRUE}
    
    ```

* This will execute the R script, defining the `wordlist` variable:

    ```{r cache = TRUE}
    source('wordlist.R')
    ```

* Here are the first elements of the `wordlist`:

    ```{r}
    head(wordlist, 20)
    ```

## Fit the chain

* This takes some time. It took about $??$ minutes on my machine. (start: 17:28)

    ```{r cache = TRUE}
    fit <- markovchainFit(
      wordlist,
      name = 'Brazilian dictionary',
      confint = FALSE
    )
    
    fit
    ```


## Answer some questions

### Letters that begin words


### Letters that end words


### Consecutive vowels


### Consecutive consonants

